Liquid Amortization - Proving amortized complexity with LiquidHaskell (Functional Pearl)

### 2021-05-21 paper/main.tex
Finish first paper draft

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
TODO
\end{abstract}

@@

\begin{itemize}
\item{Worst case is often too pessimistic}
\item{Idea: spread out expensive operations over multiple cheap ones}
\item{Usually only talks about sequence of operations without sharing}
\item{Formalization: Physicist's method}
\end{itemize}

@@

As described by Vazou et. al.~\cite{tpfa}, one can use LiquidHaskell to do complex equational reasoning that is machine checked by the compiler.

\begin{itemize}
\item{Uses refinement types}
\item{Refinments are solved by external SMT solver (z3)}
\item{With the refinement one can define proof combinators}
\item{Use equational reasoning and induction to prove interesing properties}
\end{itemize}

@@

\subsection{Stack with Multipop}

One of the simplest data structures that have an amortized complexity is a stack with two operations: Push and Multipop. $push(e)$ adds a singular element $e$ on top of the stack, while $multipop(k)$ removes the top $k$ elements. This data structure and its operations could be defined like this in Haskell. We already add some refinements to assert that $k$ is not negative.

@@

If we look at this definition we can easily see that \texttt{push} takes constant time. No matter how large the rest of the stack is, it will only allocate a new \texttt{Pushed} cell that contains a reference to the rest of the stack. Thus this operation has complexity \O{1}.

\texttt{multipop} however may pop all of the elements in the stack if $k$ is greater or equal to $n$, where $n$ is the height of the stack. So in the worst case this operation takes $n$ steps and thus it is in \O{n}.

This worst case however is a bit too pessimistic. If we pop all elements, we will not be able to do it again, as the stack is empty after the first operation. In fact, in order for an element to be popped, it first has to be pushed. So we can spread the cost for the expensive $multipop$ operation over all the constant time pushes, making \texttt{multipop} amortized \O{1}. With the banker's method this would mean for a push we could pay one time unit actual cost and one time unit extra for the bank. Then the bank has always as much time units as there are elements in the stack. Thus a \texttt{multipop} will not be able to make the bank account negative.

To formally prove this with the physicist's method, we first need to define a potential $\Phi$ and a timing function for each of our operations. The intuition for the potential is what property of the data structure makes \texttt{multipop} more expensive. Because a stack is such a simple data structure, there is basically only one property: its height. So we can define a function that calculates the height of a stack and use it as our potential. To define the timing functions, we copy the code for the operation, replace constant time steps with $1$ and adjust the recursion.

@@

First we will prove the complexity of \texttt{push}. While the amortized cost is the same as the actual cost, we need to make sure that our potential function does not cause those two to diverge. As LiquidHaskell does not allow quantifiers to ensure termination of the SMT solver, we need to guess an upper limit that our amortized runtime will have. As seen earlier, we can pay two time units for a push to amortize the pops, so we will use $2$ as our guess. To specify our proof goal, we need to name the parameters in the LiquidHaskell annotation. After than we can use equational reasoning to proof our goal.

@@

We omit the proof cases for one to three digits, as they work exactly as the other two base cases. The interesting case is when we have four digits. In this case we need to use induction to proof our goal.

@@

\section{Limitations of this approach}\label{sec:limitations}

@@

>>>>

\usepackage{cancel}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Proofs of complexity of production code is usually done in external theorem provers. This makes it necessary to translate the data structures to be proved from the production language into the language of the prover. Such a translation step may introduce errors by a mismatch in features in the two languages. We show how to prove amortized complexity of data structures directly in Haskell using LiquidHaskell~\cite{liquidhaskell}. Besides skipping the translation step, our approach also has a didactic advantage as new learners do not need to learn a new language for their proofs and can focus on the new concepts only. We do not assume prior knowledge of amortized complexity as we explain the concepts and apply them in our first case study, a simple stack with multipop. The usefulness of our approach in a production setting is verified by the second case study, the finger tree which powers \texttt{Data.Seq} from the \textit{containers} package which is widely depended on. Finally we discuss the current limitations of LiquidHaskell in this area and how they might be lifted.
\end{abstract}

@@

It is common to check how many time units an operation on a data structure takes in relation to the number of elements in the data structure. As the performance can vary wildly for the same operation depending on the state of data structure --- a linear search takes $1$ time unit if the searched element is at the front, while it takes $n$ time units if it is the last element --- usually one uses the worst case for an operation as a metric and uses Landau notation to specify it. With our linear search example we would say that it is in \O{n} because it will not perform worse than $n$ time units.

But often this worst case is too pessimistic. When an operation changes the data structure, for example by removing elements, consecutive uses of operation may be cheaper. Or to use an expensive operation it might be necessary to have a certain amount of cheap operations. In this case one uses \textit{amortized complexity} which does not only look at a single operation in isolation, but on chains of operations. In particular, it assumes that that the intermediate data structures are not shared, i.e. they are not persisted. There are multiple methods to proof amortized complexity.

As a running example we will use a stack with \textit{push} and \textit{multipop} operations. The former allocates a new element and a references to the rest of the stack and thus taking a constant amount of time. The latter removes the first $k$ elements from the stack. In the worst case, we remove all the elements, making \texttt{multipop} \O{n}. One can already see that this is too pessimistic, as after removing all elements from the stack, the next \texttt{multipop} will do nothing. In fact, every element we remove needs to be added to the stack at some point, so we can spread out the cost of \texttt{multipop} over many pushes. There are multiple techniques to prove amortized complexity, we present the two most popular ones.

\subsection{Banker's method}\label{sec:banker}

The banker's method works by having an bank account that can store time units. Operations have to pay their actual cost and may additionally pay extra into the account. Expensive operations can then use the stored time units in the bank to pay for themselves. In our example a \texttt{push} has to pay one actual time unit and it will also pay one extra time unit into the bank account. This means that the bank account always contains as many time units as there are elements on the stack. That way, any call to \texttt{multipop} can use the bank account to pay for itself and the account will never become negative.

\subsection{Physicist's method}\label{sec:physicist}

The banker's method is hard to use for a formal proof because the bank account acts as state between the operations. A simpler method is the physicist's method that does not need such an account. The idea is to define a potential $\Phi$ ("phi") for the data structure. This potential should be $0$ for an empty data structure and $\ge 0$ for all other states. Intuitively the potential will declare how much time needs to be saved up to pay for expensive operations. We then define the amortized cost of an operation as $c + \Phi(h') - \Phi(h)$ where $c$ is the actual cost for the operation, $h$ is the data structure before the operation and $h'$ is the data structure afterwards. Looking at a chain of operations $1, 2, ..., n$, the sum of this amortized time is an overestimation of the actual cost by $\Phi(h_n)$ --- because $\Phi(h_0)$ is by definition $0$ --- and thus an upper bound on the actual cost (see equation~\ref{eq:sum}).

\begin{gather}
\nonumber (c_1 + \cancel{\Phi(h_1)} - \Phi(h_0)) + (c_2 + \cancel{\Phi(h_2)} - \cancel{\Phi(h_1)}) \\
\nonumber    + ... + (c_n + \Phi(h_n) - \cancel{\Phi(h_{n - 1})}) = \\
  c_1 + c_2 + ... + c_n + \Phi(h_n) - \Phi(h_0) \label{eq:sum}
\end{gather}

For our stack we can define the potential to be the height of the stack. This fulfils the requirements as the empty stack has a potential of zero and all other states have a potential greater or equal to zero. For the actual proof of this example see the first case study in section~\ref{sec:stack}.

@@

As described by Vazou et. al.~\cite{tpfa}, one can use LiquidHaskell to do complex equational reasoning that is machine checked by the compiler. It uses refinement types~\cite{refinement_types} that add additional constraints on top of normal Haskell data types which are checked by an external SMT solver. By default LiquidHaskell uses the z3~\cite{z3} solver, but it also supports cvc4~\cite{cvc4}. For example we can declare the type of integers greater or equal to zero with a type synonym. LiquidHaskell annotions use comments with an @ sign, so that the code still compiles even without LiquidHaskell. It also allows to name arguments, so they can be referred to later in the refinement. At the moment to define functions with refinement, one has to copy the Haskell type signature into a LiquidHaskell comment and add the refinements there. A future version that is more deeply integrated into the Haskell compiler may not need this code duplication.

\begin{lstlisting}
{-@ type Nat = { x:Int | x >= 0 } @-}
\end{lstlisting}

In LiquidHaskell, all functions need to terminate. Usually a suitable termination metric is automatically guessed, but in some more complicated cases, we can also manually supply this metric (see section~\ref{sec:append} for an example of this). A simple recursive function can be marked as \textit{measure} which allows the use of the function in those termination metrics as well as in the refinement itself. With the termination, LiquidHaskell is able to generate an induction principle for the function that can be used in a proof. Because LiquidHaskell is just Haskell, this induction amounts to a recursive call to the proof and is also used when defining ordinary Haskell functions with refinement on their types. For example the fact that the length of a list is never negative.

\begin{lstlisting}
{-@ measure length @-}
{-@ length :: [a] -> Nat @-}
length :: [a] -> Int
length [] = 0 -- trivial, 0 >= 0
length (x:xs) = 1 + length xs
    -- by induction: length xs >= 0
    -- so 1 + length xs >= 0
\end{lstlisting}

LiquidHaskell also provides a set of proof combinators and a \texttt{Proof} type. This type is just a type alias for the normal Haskell unit type. The reasoning comes from the \texttt{===} operator that asserts that both sides of the equality are truly equal. As a simple example we can proof that the length of an \texttt{append} function is the sum of the lengths of the two original lists. As append is not a \textit{measure} (it does not return a natural number), we have to use \textit{reflection}~\cite{reflection} to lift it to the logic level of LiquidHaskell. The proof itself is done by case distinction on the first argument. This distinction is done by pattern matching on it like one would in a normal Haskell function. We start with stating the left hand side of the equation how it is in the refinement type and try to get to the right hand side of the equality by simplification and induction.

\begin{lstlisting}
{-@ reflect append @-}
append :: [a] -> [a] -> [a]
append [] ys = ys
append (x:xs) ys = x:(append xs ys)

{-@ lengthP :: xs:[a] -> ys:[a] ->
  { length (append xs ys) ==
      length xs + length ys }
@-}
lengthP :: [a] -> [a] -> Proof

-- base case
lengthP [] ys = length (append [] ys)
  -- Use definition of append [] _
    === length ys
    === 0 + length ys
  -- use definition of length []
    === length [] + length ys
    *** QED

-- recursive case
lengthP (x:xs) ys =
    length (append (x:xs) ys)
  -- Use definition of append
    === length (x:(append xs ys))
  -- Use definition of length
    === 1 + length (append xs ys)
  -- Use induction with the ? operator
    ? lengthP xs ys
    === 1 + length xs + length ys
  -- Use definition of length
    === length (x:xs) + length ys
    *** QED
\end{lstlisting}

The triple star operator (\texttt{***}) and the \texttt{QED} type at the end are just to cast our proof to Haskell's unit type and to make the proof look nicer.

@@

\subsection{Stack with Multipop}\label{sec:stack}

To continue with the example from section~\ref{sec:complexity}, we want to prove that $multipop$ has an amortized complexity of \O{1}. This data structure and its operations could be defined like this in Haskell. We already add some refinements to assert that $k$ is not negative.

@@

To formally prove this with the physicist's method, we first need to define a potential $\Phi$ and a timing function for each of our operations that gives us the actual cost of that operation. As seen in section~\ref{sec:physicist}, we use the height of the stack as our potential. Thus we need to define a function that calculates the height of a stack. To define the timing functions, we copy the code for the operation, replace constant time steps with $1$ and adjust the recursion.

@@

First we will prove the complexity of \texttt{push}. While the amortized cost is the same as the actual cost, we need to make sure that our potential function does not cause those two to diverge. As LiquidHaskell does not allow quantifiers to ensure termination of the SMT solver, we need to guess an upper limit that our amortized runtime will have. As seen earlier in section~\ref{sec:banker}, we can pay two time units for a push to amortize the pops, so we will use $2$ as our guess. To specify our proof goal, we need to name the parameters in the LiquidHaskell annotation. After than we can use equational reasoning to proof our goal.

@@

We omit the proof cases for one to three digits, as they work exactly as the other two base cases. The constant however will not change in the other base cases, so $3$ really is our upper bound to use here. The interesting case is when we have four digits. In this case we need to use induction to proof our goal.

@@

@@

\section{Full proof code for the stack}

data Stack a = Empty | Elem a (Stack a)

{-@ reflect push @-}
push :: a -> Stack a -> Stack a
push = Elem

{-@ multipop :: Nat -> Stack a -> Stack a @-}
multipop :: Int -> Stack a -> Stack a
multipop _ Empty = Empty
multipop 0 s = s
multipop n (Elem _ s) = multipop (n - 1) s

{-@ reflect phi @-}
{-@ phi :: Stack a -> Nat @-}
phi :: Stack a -> Int
phi Empty = 0
phi (Elem _ s) = 1 + phi s

{-@ reflect pushT @-}
{-@ pushT :: a -> Stack a -> Nat @-}
pushT :: a -> Stack a -> Int
pushT _ _ = 1

{-@ reflect multipopT @-}
{-@ multipopT :: Nat -> Stack a -> Nat @-}
multipopT :: Int -> Stack a -> Int
multipopT _ Empty = 1
multipopT 0 _ = 1
multipopT n (Elem _ s) = 1 + multipopT (n - 1) s

{-@ pushP :: x:a -> s:Stack a ->
  { pushT x s + phi (push x s) - phi s <= 2 }
@-}
pushP :: a -> Stack a -> Proof
pushP x s = pushT x s + phi (push x s) - phi s
-- Use definition of push and pushT
  === 1 + phi (Elem x s) - phi s
-- Use definition of phi (Elem _ _)
  === 1 + (1 + phi s) - phi s
-- phi s gets canceled, 1 + 1 <= 2
  *** QED

{-@ multipopP :: k:Nat -> s:Stack a ->
  { multipopT k s + phi (multipop k s)
      - phi s <= 2 }
@-}
multipopP :: Int -> Stack a -> Proof
multipopP 0 s = multipopT 0 s
    + phi (multipop 0 s) - phi s
-- Use definition of multipop/T 0 _
  === 1 + phi s - phi s
-- phi s cancels, 1 <= 2
  *** QED
multipopP k Empty = multipopT k Empty
    + phi (multipop k Empty) - phi Empty
-- Use definition of multipop/T _ Empty
  === 1 + phi Empty - phi Empty
-- phi Empty cancels, 1 <= 2
  *** QED
multipopP k xs@(Elem x s) = multipopT k xs
    + phi (multipop k xs) - phi xs
-- Use definition of multipopT for recusive case
  === (1 + multipopT (k - 1) s)
    + phi (multipop k xs) - phi xs
-- Use definition of multipop for recusive case
  === (1 + multipopT (k - 1) s)
    + phi (multipop (k - 1) s) - phi xs
-- Use definition of phi (Elem _ _)
  === (1 + multipopT (k - 1) s)
    + phi (multipop (k - 1) s) - (1 + phi s)
-- Remove parenthesis
  === 1 + multipopT (k - 1) s
    + phi (multipop (k - 1) s) - 1 - phi s
-- 1s cancel out
  === multipopT (k - 1) s
    + phi (multipop (k - 1) s) - phi s
-- use induction
  ? multipopP (k - 1) s
  *** QED
\end{lstlisting}

\section{Full proof code of the finger tree}\label{ap:fingertree}

data Digit a = One a
  | Two a a
  | Three a a a
  | Four a a a a

{-@ reflect cons @-}
cons :: a -> FingerTree a -> FingerTree a
cons a EmptyT = Single a
cons a (Single b) =
  Deep (One a) EmptyT (One b)
cons a (Deep (One b) m sf) =
  Deep (Two a b) m sf
cons a (Deep (Two b c) m sf) =
  Deep (Three a b c) m sf
cons a (Deep (Three b c d) m sf) =
  Deep (Four a b c d) m sf
cons a (Deep (Four b c d e) m sf) =
  Deep (Two a b) (cons (Node3 c d e) m) sf

{-@ reflect snoc @-}
snoc :: FingerTree a -> a -> FingerTree a
snoc EmptyT x = Single x
snoc (Single a) x =
  Deep (One a) EmptyT (One x)
snoc (Deep pr m (One a)) x =
  Deep pr m (Two a x)
snoc (Deep pr m (Two a b)) x =
  Deep pr m (Three a b x)
snoc (Deep pr m (Three a b c)) x =
  Deep pr m (Four a b c x)
snoc (Deep pr m (Four a b c d)) x =
  Deep pr (snoc m (Node3 a b c)) (Two d x)

{-@ reflect dang @-}
{-@ dang :: Digit a -> Nat @-}
dang :: Digit a -> Int
dang One{} = 1
dang Two{} = 0
dang Three{} = 0
dang Four{} = 1

{-@ reflect phi @-}
{-@ phi :: FingerTree a -> Nat @-}
phi :: FingerTree a -> Int
phi EmptyT = 0
phi Single{} = 0
phi (Deep pr m sf) =
  dang pr + phi m + dang sf

{-@ reflect consT @-}
{-@ consT :: a -> FingerTree a -> Nat @-}
consT :: a -> FingerTree a -> Int
consT _ EmptyT = 1
consT _ Single{} = 1
consT _ (Deep One{} _ _) = 1
consT _ (Deep Two{} _ _) = 1
consT _ (Deep Three{} _ _) = 1
consT _ (Deep (Four _ c d e) m _) =
  1 + consT (Node3 c d e) m

{-@ reflect snocT @-}
{-@ snocT :: FingerTree a -> a -> Nat @-}
snocT :: FingerTree a -> a -> Int
snocT EmptyT _ = 1
snocT Single{} _ = 1
snocT (Deep _ _ One{}) _ = 1
snocT (Deep _ _ Two{}) _ = 1
snocT (Deep _ _ Three{}) _ = 1
snocT (Deep _ m (Four a b c _)) _ =
    1 + snocT m (Node3 a b c)

{-@ consP :: x:a -> t:FingerTree a ->
  { consT x t + phi (cons x t) - phi t <= 3 }
@-}
consP :: a -> FingerTree a -> Proof
consP x t@EmptyT =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T _ EmptyT
  === 1 + phi (Single x) - phi EmptyT
-- Use definition of phi
  === 1 + 0 - 0 -- trivial, 1 <= 3
  *** QED
consP x t@(Single a) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T _ (Single a)
  === 1 + phi (Deep (One x) EmptyT (One a))
        - phi (Single a)
-- Use definition of phi
  === 1 + dang (One x) + phi EmptyT
        + dang (One a) - 0
-- Use definition of phi and dang
  === 1 + 1 + 0 + 1 - 0 -- trivial, 3 <= 3
  *** QED
consP x t@(Deep (One a) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T
  === 1 + phi (Deep (Two x a) m sf) - phi t
-- Use definition of phi
  === 1 + dang (Two x a) + phi m + dang sf
    - (dang (One a) + phi m + dang sf)
-- Remove parenthesis, use definition of dang
  === 1 + 0 + phi m + dang sf
    - 1 - phi m - dang sf
-- everything cancels
  === 0 -- trivial, 0 <= 3
  *** QED
consP x t@(Deep (Two a b) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T
  === 1 + phi (Deep (Three x a b) m sf) - phi t
-- Use definition of phi
  === 1 + dang (Three x a b) + phi m + dang sf
    - (dang (Two a b) + phi m + dang sf)
-- Remove parenthesis, use definition of dang
  === 1 + 0 + phi m + dang sf
    - 0 - phi m - dang sf
-- phi and dang cancels
  === 1 -- trivial, 1 <= 3
  *** QED
consP x t@(Deep (Three a b c) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T
  === 1 + phi (Deep (Four x a b c) m sf) - phi t
-- Use definition of phi
  === 1 + dang (Four x a b c) + phi m + dang sf
    - (dang (Three a b c) + phi m + dang sf)
-- Remove parenthesis, use definition of dang
  === 1 + 1 + phi m + dang sf
    - 0 - phi m - dang sf
-- phi and dang cancels
  === 2 -- trivial, 2 <= 3
  *** QED
consP x t@(Deep (Four a b c d) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of consT
  === (1 + consT (Node3 b c d) m)
  + phi (cons x t) - phi t
-- Use definition of cons
  === (1 + consT (Node3 b c d) m)
  + phi (Deep (Two x a) (cons (Node3 b c d) m) sf)
  - phi t
-- Use definition of phi
  === (1 + consT (Node3 b c d) m)
  + dang (Two x a) + phi (cons (Node3 b c d) m)
  + dang sf
  - (dang (Four a b c d) + phi m + dang sf)
-- Remove parenthesis
  === 1 + consT (Node3 b c d) m
  + dang (Two x a) + phi (cons (Node3 b c d) m)
  + dang sf
  - dang (Four a b c d) - phi m - dang sf
-- dang sf cancels, use definition of dang
  === 1 + consT (Node3 b c d) m
  + 0 + phi (cons (Node3 b c d) m) - 1 - phi m
-- 1s cancel
  === consT (Node3 b c d) m
  + phi (cons (Node3 b c d) m) - phi m
-- Use induction
  ? consP (Node3 b c d) m
  *** QED

{-@ snocP :: t:FingerTree a -> x:a ->
  { snocT t x + phi (snoc t x) - phi t <= 3 }
@-}
snocP :: a -> FingerTree a -> Proof
snocP t@EmptyT x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T EmptyT _
  === 1 + phi (Single x) - phi EmptyT
-- Use definition of phi
  === 1 + 0 - 0 -- trivial, 1 <= 3
  *** QED
snocP t@(Single a) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T (Single a) _
  === 1 + phi (Deep (One a) EmptyT (One x))
        - phi (Single a)
-- Use definition of phi
  === 1 + dang (One a) + phi EmptyT
        + dang (One x) - 0
-- Use definition of phi and dang
  === 1 + 1 + 0 + 1 - 0 -- trivial, 3 <= 3
  *** QED
snocP t@(Deep pr m (One a)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T
  === 1 + phi (Deep pr m (Two a x)) - phi t
-- Use definition of phi
  === 1 + dang pr + phi m + dang (Two a x)
    - (dang pr + phi m + dang (One a))
-- Remove parenthesis, use definition of dang
  === 1 + dang pr + phi m + 0
    - dang pr - phi m - 1
-- everything cancels
  === 0 -- trivial, 0 <= 3
  *** QED
snocP t@(Deep pr m (Two a b)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T
  === 1 + phi (Deep pr m (Three a b x)) - phi t
-- Use definition of phi
  === 1 + dang pr + phi m + dang (Three a b x)
    - (dang pr + phi m + dang (Two a b))
-- Remove parenthesis, use definition of dang
  === 1 + dang pr + phi m + 0
    - dang pr - phi m - 0
-- phi and dang cancels
  === 1 -- trivial, 1 <= 3
  *** QED
snocP t@(Deep pr m (Three a b c)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T
  === 1 + phi (Deep pr m (Four a b c x)) - phi t
-- Use definition of phi
  === 1 + dang pr + phi m + dang (Four a b c x)
    - (dang pr + phi m + dang (Three a b c))
-- Remove parenthesis, use definition of dang
  === 1 + dang pr + phi m + 1
    - dang pr - phi m - 0
-- phi and dang cancels
  === 2 -- trivial, 2 <= 3
  *** QED
snocP t@(Deep pr m (Four a b c d)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snocT
  === (1 + snocT m (Node3 a b c))
  + phi (snoc t x) - phi t
-- Use definition of snoc
  === (1 + snocT m (Node3 a b c))
  + phi (Deep pr (snoc m (Node3 a b c)) (Two x a))
  - phi t
-- Use definition of phi
  === (1 + snocT m (Node3 a b c))
  + dang pr + phi (snoc m (Node3 a b c))
  + dang (Two x a)
  - (dang pr + phi m + dang (Four a b c d))
-- Remove parenthesis
  === 1 + snocT m (Node3 a b c)
  + dang pr + phi (snoc m (Node3 a b c))
  + dang (Two x a)
  - dang pr - phi m - dang (Four a b c d))
-- dang pr cancels, use definition of dang
  === 1 + snocT m (Node3 a b c)
  + phi (snoc m (Node3 a b c)) + 0 - phi m - 1
-- 1s cancel
  === snocT m (Node3 a b c)
  + phi (snoc m (Node3 a b c)) - phi m
-- Use induction
  ? snocP m (Node3 a b c)
  *** QED

\end{lstlisting}
